Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-09-10 documents/ovando-etal-assessing-global-fisheries.Rmd
run on squid

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::pdf_document2: default
  bookdown::word_document2:
    reference_docx: word-template.docx
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that catch data by themselves cannot provide reliable estimates of fishery health. Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

@@

@@

# rishi notes

* more assertive on what the data needs really are
  * where to apply globally 
* For places with less coverage, what would we actually recommend there 
* what do we know about catchabiliy
- e.g. 51, fleets became industrial 
  - hammer down on that's the issue with the effort 
  - how do we translate effort into f bey area
  - here's

Target journal: Nature Sustainability

https://www.nature.com/natsustain/about/content

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished ,and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].

Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (@free2020 and references therein). These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, given that a global assessment by its nature requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks (high variance, low bias).Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status (B/B~MSY~). These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.

xx spend time on this tomorrow xx
Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
               booktabs = TRUE) %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

@@

xx move some of this to intro and chop downXX
We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model. `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.

<!-- ## Case Studies  -->

<!-- Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status. For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region. We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index. By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic). Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17. Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).  -->

 <!-- Evaluating the validity of data-limited assessments methods can be challenging. Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data. We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments. However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.  -->

 <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region. As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx. We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.  -->

<!-- If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense. However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal. Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar"). While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.  -->

<!-- ```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio. Each point is a stock."} -->
<!-- ei_kobe_plot  -->
<!-- ``` -->

We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx). Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since. Current catches in Europe are much higher than those at the start of the catch history. Catches in northeast Asia have steadily increased over time, with some stabilization as of late. For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high. However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time. As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status. Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions. Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe. However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).

XX add stock assessment estimates for both regions? XX Maybe just do matching stocks?

```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019 Each point is a stock, point size is a function of stock size."}
continent_kobe_plot
```

## Conflicting Estimates of Global Status

```{r perf-tab}
breaks <- c(0.8, 1.2)

knitr::kable(perf, digits = 2,
             caption = "Global summary statistics of models using different sources of data. mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).",
              format = "pipe")

@@

@@

## Example Use of `sraplus`

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess. Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

@@

  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))

breaks <- c(0.8, 1.2)

fao_area_codes <- fao %>% 
  ungroup() %>% 
  select(fao_area, fao_area_code) %>% 
  unique()

@@

<!-- # rishi notes -->

<!-- * more assertive on what the data needs really are -->
<!--   * where to apply globally  -->
<!-- * For places with less coverage, what would we actually recommend there  -->
<!-- * what do we know about catchabiliy -->
<!-- - e.g. 51, fleets became industrial  -->
<!--   - hammer down on that's the issue with the effort  -->
<!--   - how do we translate effort into f bey area -->
<!--   - here's  -->

# Publication Requirements

Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content)

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal depends on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking the global state of fisheries. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed). Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].

Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks. Each of the published catch-based methods reviewed in Free et al. 2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Examining stock assessment data from RAM from the  Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020]. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)). While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,

@@

o_by_a <- sofia_status %>% 
  left_join(fao_area_codes, by = "fao_area_code") %>% 
  group_by(fao_area) %>% 
  summarise(`FAO % Overfished` = percent(mean(status == "O")))

ss <-
  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
                    sheet = "state_space_results_08_28_2018") %>%
  janitor::clean_names()

ss_lookup <-
  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
                    sheet = "ss_to_fao") %>%
  janitor::clean_names() %>% 
  mutate(tmp = map2(region, fao_area_codes, ~data.frame(region = .x, fao_area_code = as.integer(str_split(.y, pattern = ',',simplify = TRUE))))) %>% 
  select(tmp) %>% 
  unnest(cols = tmp) %>% 
    left_join(fao_area_codes, by = "fao_area_code")

state_space_results <- ss_lookup %>% 
  left_join(ss, by = "region") %>% 
  filter(variable == "BvB") %>% 
  group_by(fao_area) %>% 
  filter(year == max(year)) %>% 
  summarise(`B/BMSY RAM` = round(mean(dlm_geomean),2)) %>% 
  ungroup()

# ram_by_region %>% 
#   ggplot(aes(b_v_bmsy)) +
#   geom_histogram() + 
#   facet_wrap(~fao_area)

readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries? We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management through Fisheries Management Index (FMI) scores [@melnychuk2017], trawl-fishing footprints through swept-area-ratio (SAR) values [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model.

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data (combinations of catches, effort, fisheries management index, and swept area ratio values) relative values reported in the RAM database.

 <!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.  -->

XX discussion with Jim. How can we frame this. I would actually push back the idea that we have learned about stock status. I buy that the catch data can update your prior on growth parameters, but by the rules we're playing with here, it shouldn't update our prior on stock status should it. Now, the only thing I can't wrap my head aroudn is the case where catches are all zero, and you have a prior that the stock is collapsed. I guess my question is from a strikely Bayes theorem perspective, where in here is your prior on stock status being updated, since there's no likelihood as written, except for the binomial process that says that the stock exists.

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. It is critical to note that in both of these cases the model has learned nothing new about recent stock status that was not encoded in the supplied priors, catch data, and functional form of the model; `sraplus` simply helps reveal our implicit priors on other parameters (e.g. maximum sustainable yield, MSY), and provides timeseries of measures such as B/B~MSY~. That, in the absence of data to actually fit to such as an index of abundance, methods such as CMSY or sraplus can learn anything new about a stock's status that is not already implicitly coded in the priors, catch data, or specified functional relationships, is a common misconception (see [Methods](#Methods) a for detailed explanation of why this is). Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values. An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).

```{r cs-plot, fig.cap="Estimates of stock status for case study fisheries using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index and Swept Area Ratio scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%. Each point is a stock, point size is a function of stock size. The shaded hexagons in the background show the distribution of stock status estimates from the RAM Legacy Stock Assessment Database for each of the stocks included in this figure, with darker colors representing increased concentrations of RAM stocks."}

ex_kobe_plot
```

That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.

While SAR and FMI data are available for a wide number of stocks, they are not ubiquitous enough to provide stock-specific priors for the number of fisheries needed for a global assessment. In addition, predicting stock status based on SAR, FMI, or catch patterns requires that the same patterns that exist within the training data as exist in other stocks to be used for prediction, which is far from clear given the highly managed nature of the stocks for which we have reliable estimates of stock status. We can extend the ideas presented in the case study, using combinations of catch, SAR, FMI, and effort data to produce estimates of stock status at the global level.

As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data. For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock. For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level). This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.

As a proof of concept, we first estimated B/B~MSY~ of our candidate RAM stocks by using `sraplus` to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations of the regional-level data made available to them (Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

```{r dat-desc}
dat_description <- tribble(~`Data Name`, ~Description,
        "RAM-Data", "Fit to abundance index from RAM",
        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "catch-only", "priors on terminal B/Bmsy informed by catch history",
        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")

Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality. Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM. Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.

Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves. These becomes clearer once we consider the performance of `sraplus` models fit to combinations of our broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])` (since fisheries are not uniformly distributed across each bin).

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. Across the models performance (or lack thereof) is not consistent in space: use of different data performed best or worst for different FAO regions. For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.

```{r perf-tab}

knitr::kable(perf, digits = 2,
             caption = "Global performance statistics in the most recent year available of models using different sources of data. mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished). Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
              format = "pipe")

@@

<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->

<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->

@@
